Automatically proving equivalence by type-safe reflection

### 2016-01-07 paper-JFP-DTP/content/basicIdeas.tex
paper JFP DTP continued

Now, we need to define the function reduce. To do that, we have to decide a canonical representation of associative natural numbers. We decide that the left associative form will be the canonical representation. Thus, the $reduce$ function has to rewrite the metaified term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ... + xn)$. To do so, one possibility is to define a new datatype which captures this property, and to write a function going from $Expr$ to this new type. Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.

@@

We notice that for transforming the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities. \\
\\
We've got two metavariables to prove. The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ \Gamma)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.

@@

We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively.

>>>>

Now, we need to define the function reduce. To do that, we have to decide a canonical representation of associative natural numbers. We decide that the left associative form will be the canonical representation. Thus, the $reduce$ function has to rewrite the reflected term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ... + xn)$. To do so, one possibility is to define a new datatype which captures this property, and to write a function going from $Expr$ to this new type. Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.

@@

We notice that in order to transform the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities. \\
\\
We've got two metavariables to prove. The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ \Gamma)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.

@@

We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively. 
By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values. That implies that when applying this function to some arguments, the arguments aren't reduce before the evaluation of this function.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}

%reflection
reflectList : (G : List Nat) ->
          (x : Nat) -> (G' ** Expr (G ++ G') x)
reflectList G Z = ([] ** Zero)

...

\end{verbatim}
\end{center}
\caption{Reflection for natural numbers}
\figrule
\end{figure}
